---
title: "Making maps of Italy with R"
author: "Cynthia Anstey"
date: "`r Sys.Date()`"
output:
  html_document:
---


### dati sorigente
###### italiano
Dati sorgente
<p>I file contenuti in questo archivio sono derivati dai dati di OpenStreetMap e sono distribuiti sotto la Licenza Open Database 1.0. Per maggiori informazioni sul progetto, visita www.openstreetmap.org.</p> <p>Questo file contiene dati di OpenStreetMap aggiornati al 2024-08-23T20:21:17Z. Ogni giorno verrà resa disponibile una nuova versione di questo file al seguente link:</p> <p><a href="http://download.geofabrik.de/europe/italy/centro-latest-free.shp.zip">http://download.geofabrik.de/europe/italy/centro-latest-free.shp.zip</a></p> 
<p>Una documentazione sui livelli di questo shapefile è disponibile qui:</p> <p><a href="http://download.geofabrik.de/osm-data-in-gis-formats-free.pdf">http://download.geofabrik.de/osm-data-in-gis-formats-free.pdf</a></p> 
<p>Ci sono 17 file .shp all'interno del file zip scaricato al 27-08-2024. Ho selezionato un file che si chiama "places" e "free", sperando che la struttura fosse adatta per esercitarmi.</p>

### source data
###### english
The files in this archive are derived from OpenStreetMap data and are licensed under the Open Database 1.0 License. For more details, visit [OpenStreetMap](www.openstreetmap.org).

- **File link**: [Download Shapefile](http://download.geofabrik.de/europe/italy/centro-latest-free.shp.zip)
- **Documentation**: [Layer Documentation](http://download.geofabrik.de/osm-data-in-gis-formats-free.pdf)


### resources
<p>Cookbooks that helped the learning process.</p> 
- [ggplot2 and sf integration](https://r-spatial.org/r/2018/10/25/ggplot2-sf.html)
- [Shapefile](https://r-graphics.org/recipe-miscgraph-map-shapefile)
- [advanced mapping](https://bookdown.org/robinlovelace/geocompr/adv-map.html)


Install Libraries

```{r}
library("ggplot2")
theme_set(theme_bw())
library("sf")
```

```{r}
library("rnaturalearth")
library("rnaturalearthdata")
```


```{r}
library(tmap)
```


```{r}
library(maps)
library(mapdata)
library(mapview)
library(dplyr)
library(magrittr)
library(leaflet)
```

```{r}
#After importing a file, use st_read to read it into R.
italy_shp <- st_read("C:\\Users\\Public\\Downloads\\centro-latest-free.shp\\gis_osm_places_free_1.shp")
```


```{r}
italy_shp1 <- st_read("C:\\Users\\Public\\Downloads\\centro-latest-free.shp\\gis_osm_places_a_free_1.shp")
```


Use ggplot give the file a geometric simple form (sf). italy_shp is plotted first then italy_shp1. 
```{r}
ggplot(italy_shp)+
geom_sf()
```
```{r}
ggplot(italy_shp1)+
geom_sf()
```
The following code layers both spatial plots. The selected file, labeled as "Centro," includes geometric points that indicate locations considered central in Italy.

```{r}
ggplot() +
  geom_sf(data = italy_shp, fill = "transparent", color = "lightblue", size = 0.5) +
  geom_sf(data = italy_shp1, fill = "transparent", color = "darkorange", size = 0.5) +
  theme_minimal() +
  labs(title = "Layered Spatial Plots",
       subtitle = "Overlaying italy_shp and italy_shp1",
       caption = "Source: Example Data") +
  theme(legend.position = "bottom")
```



Inspect the Data. 
```{r}
summary(italy_shp)
```

Read name as a list.
```{r}
names(italy_shp)
```

Read the file structure.
```{r}
str(italy_shp)
```

Create a Tibble.
```{r}
tibble(italy_shp)
```

Get some basic statistics.
```{r}
mean_population <- mean(italy_shp$population)
median_population <- median(italy_shp$population)
sd_population <- min(italy_shp$population)
min_population <- max(italy_shp$population)
mean_population
median_population
sd_population
min_population
```

Create a summary of the population column.
```{r}
summary(italy_shp$population)
```

Create bins for population and make a map by pop bins.
```{r}
italy_shp_bins <- italy_shp %>%
  mutate(population_bin = cut(population, breaks = c(0, 1, 500, 1000, 2000, 4000, 8000, 10000, 12000, 15000, 18000, 20000, 22000, 30000, 60000, Inf), 
                               labels = c("0", "1-500", "500-1000", "1000-2000", "2000-4000", "4000-8000", "8000-10000","10000-12000", "12000-15000","15000-18000", "18000-20000", "20000-22000", "22000-30000", "30000-60000", "60000 +")))

# Count the number of towns in each population bin
population_bins <- italy_shp_bins %>%
  count(population_bin)

# Print distribution
print(population_bins)
```

Plotting Population Histograms and Bar Charts
```{r}
italy_shp_histogram <- italy_shp_bins %>%
  mutate(population_bin = cut(population,
                              breaks = c(0, 1000, 5000, 10000, 50000, 80000, 100000, Inf),
                              labels = c("0-1k", "1k-5k", "5k-10k", "10k-50k", "50k-80k", "80k-100k", "100k+")))

ggplot(italy_shp_histogram, aes(x = population_bin)) +
  geom_bar(fill = "lightblue", color = "black", width = 0.4) +  # Adjust width here
  coord_flip() +
  labs(
    x = "Population Range",
    y = "Count",
    title = "Count of Entities by Population Range"
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 8, face = "bold"),
    axis.title.y = element_text(size = 8, face = "bold"),
    axis.text = element_text(size = 9),
    plot.title = element_text(size = 9, face = "bold")
  ) +
  geom_text(stat = "count", aes(label = ..count..), hjust = -0.2, size = 2.0, color = "black")  # Adjust size here
options(repr.plot.width=15, repr.plot.height=6)
```

```{r}
italy_shp$fclass <- as.character(italy_shp$fclass)
```

```{r}
# Check specific columns
table(italy_shp$fclass)
```

 
Sorting Names.                                                                                                                       
```{r}

# Sort names
sorted_names <- sort(italy_shp$name)

# Display the first 10 names
print(head(sorted_names, 10))

# Or display a random sample of 10 names
set.seed(123)  # Set seed for reproducibility
sampled_names <- sample(sorted_names, 10)
print(sampled_names)

# Or display the first 10 unique names
unique_names <- unique(italy_shp$name)
print(head(unique_names, 10))
```
Cleaning Names.
```{r}
library(stringi)
italy_shp$name_clean <- italy_shp$name %>%
# Remove escaped double quotes (\")
gsub(pattern = "\\\\\"", replacement = "", .) %>%
# Remove extra double quotes (")
gsub(pattern = "\"", replacement = "", .) %>%
# Remove backslashes
gsub(pattern = "\\\\", replacement = "", .) %>%
# Remove single quotes
gsub(pattern = "'", replacement = "", .) %>%
# Remove accent marks
stri_trans_general(id = "Latin-ASCII")

# Check the results
# Sort names
sorted_names <- sort(italy_shp$name_clean)

# Display the first 50 names
print(head(sorted_names, 50))
```
Additional inspection.
```{r}
# Inspect the data
head(italy_shp$name_clean)
str(italy_shp$name_clean)
str(italy_shp$geometry)
```

```{r}
# Count occurrences of each name
duplicates <- italy_shp %>%
  count(name_clean) %>%
  filter(n > 1)  # Filter for names that appear more than once

print(duplicates)
```
```{r}
library(mapview)
mapview(italy_shp)
```


Grouping by fclass and Counting Entities

```{r}
# Group by 'fclass' and summarize
df_grouped <- italy_shp %>%
  group_by(fclass) %>%
  summarise(
    count = n(),  # Count of entities in each fclass
    # Combine all names into a single string per group
    entities = paste(name_clean, collapse = ", "),
    # Combine all geometries into a single geometry per group
    geometry = st_union(geometry),
    .groups = 'drop'  # Drop the grouping structure in the result
  )

# Print the grouped summary
print(df_grouped)
```
```{r}
# Convert df_grouped to an sf object
italy_shp<- st_as_sf(italy_shp)
```


```{r}
# Plot using mapview
mapview(italy_shp, zcol = "fclass", legend = TRUE, map.types = "OpenStreetMap")
```


```{r}
top_cities <- italy_shp %>%
  
  arrange(desc(population)) %>%
  slice_head(n = 10)
top_cities <- top_cities %>%
  mutate(log_population = log1p(population))  # log1p is used to handle zero and small values gracefully
# Plot using mapview
mapview(top_cities, zcol = "population", legend = FALSE, map.types = "OpenStreetMap")
```
```{r}
library(xaringan)
```

